\begin{abstract}
Asymmetric multicore processors (AMP) are necessary for extracting performance in an era of limited power budget and dark silicon. Despite becoming increasingly prevalent, our software fails to use them efficiently. OS schedulers, in particular, handle asymmetry only under restricted scenarios, if at all. We have efficient symmetric schedulers, efficient asymmetric schedulers for single-threaded workloads, and efficient asymmetric schedulers for single program workloads. What we do not have is a scheduler that can handle all three factors affecting AMP scheduling: core affinity, thread criticality, and scheduling fairness.

To address this problem, this paper introduces the first general purpose asymmetry-aware scheduler targeting multi-threaded multi-programmed workloads. It estimates the performance of each thread on each type of core and it identifies communication patterns and bottleneck threads. With this information, the scheduler makes coordinated core assignment and thread selection decisions that still provide each application its fair share of the processor's time.

 We demonstrate performance gains of up to 25\% compared to Linux and up to 21\% compared to a state-of-the-art AMP-aware scheduler on multi-threaded multi-programmed workloads from 312 experiments with different configurations and programs compositions from PARSEC, SPLASH2 benchmark suites, using the GEM5 simulation of ARM big.LITTLE architectures.
     
%We evaluated our scheduler on GEM5 using an ARM big.LITTLE architecture and multi-thread multi-program workloads, mixing benchmarks from PARSEC, SPLASH2, and SPEC2006. Compared against the Linux CFS scheduler and a state-of-the-art AMP-aware scheduler, our experimental results show up to 51\% performance gain, or 5\% to 27\% on average depending on the hardware setup.
\end{abstract}


%\begin{abstract}
%Due to an excellent tradeoff between performance and energy efficiency, asymmetric multicore processors (AMPs) are becoming increasingly popular in the era of limited power and dark silicon. However, OS schedulers are still not fully adapted to them and are thus unable to exploit them efficiently. In this paper, we introduce a novel general-purpose asymmetry-aware scheduler that targets \emph{multi-threaded multi-programmed} workloads. By estimating the performance of each thread on each type of core, identifying communication patterns and bottleneck threads and making, based on this information, coordinated core assignment and thread selection decisions, it simultanously addresses the three core factors that affect AMP scheduling - \emph{core affinity, thread criticality and scheduling fairness}. We demonstrate performance gains of up to XX\% on general workloads of multithreaded applications from PARSEC, SPLASH2 and SPEC2006 benchmark suites, and up to XX\% on workloads of synchronisation-intensive benchmarks compared to the state-of-the-art Linux and XX schedulers, using the GEM5 simulation of ARM big.LITTLE architectures.
%\end{abstract}
